# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Unzip and load the data
```{r Processing, echo=TRUE}
unzip("activity.zip")
data <- read.csv(file = "activity.csv",
                 na.strings = "NA",
                 colClasses = c("numeric", "Date", "numeric"))
```


## What is mean total number of steps taken per day?
```{r Histogram, echo=TRUE}
library(ggplot2)
library(plyr)

nsteps.per.day <- ddply(data, "date", summarise, 
                        total.steps = sum(steps, na.rm = TRUE))
head(nsteps.per.day)

p <- ggplot(data = nsteps.per.day) +
        geom_histogram(aes(x = total.steps), binwidth = 2000) +
        labs(x = "Total number of steps taken each day",
             y = "Count",
             title = "Histogram of the Total Number of Steps Taken Each Day")   
print(p)

# Mean of the total number of steps.
mean(nsteps.per.day$total.steps)

# Median of the total number of steps.
median(nsteps.per.day$total.steps)
```

## What is the average daily activity pattern?

```{r TimeSeries, echo=TRUE}
ndays <- nrow(nsteps.per.day) # Total number of days

avg.nsteps.per.interval <- ddply(data, "interval", summarise, 
                                average.nsteps = sum(steps, na.rm = TRUE)/ndays)

head(avg.nsteps.per.interval)

p <- ggplot(avg.nsteps.per.interval, aes(x = interval, y = average.nsteps)) +
        geom_line() +
        labs(x = "5-minute interval",
             y = "Daily average number of steps",
             title = "Average Daily Activity Pattern")

print(p)

## 5-minute interval with the maximum number of steps on daily average
max(avg.nsteps.per.interval$average.nsteps)
```


## Imputing missing values

#### Total number of missing values
```{r NumberOfMV, echo=TRUE}
summary(data)
mv <- sum(is.na(data))

# Missing Values
print(mv)
```

#### Imputing missing values using the mean for the 5-minute interval

```{r MissingValuesImputation, echo=TRUE}
## Missing values were imputed using the average number of steps taken,
## averaged across all days. This data is contained in the avg.nsteps.per.interval
## dataframe reported above.

## The function get.avg.step returns the average number of steps taken, 
## rounded to the nearest integer, for a particular interval n.
get.avg.step <- function(n){
        round(
        avg.nsteps.per.interval$average.nsteps[avg.nsteps.per.interval$interval == n])  
}

## data.mi is a new dataset that is equal to the original dataset 
## but with the missing data filled in.

data.mi <- data
data.mi[which(is.na(data.mi)),]$steps <- sapply(
        data.mi[which(is.na(data.mi)),]$interval, get.avg.step)

summary(data.mi)
head(data.mi)
```

#### Histogram of the total number of steps taken each day after missing value imputation

```{r HistogramAfterImputation, echo=TRUE}
tsteps.df <- ddply(data.mi, "date", summarise, total.steps=sum(steps))

p <- ggplot(data = tsteps.df) +
        geom_histogram(aes(x = total.steps), binwidth = 2000) +
        labs(x = "Total number of steps taken each day",
             y = "Count",
             title = "Histogram of the Total Number of Steps Taken Each Day\nAfter Missing Data Imputation")   
print(p)

# Mean of the total number of steps after m.d. imputation
mean(tsteps.df$total.steps)

# Median of the total number of steps after m.d. imputation
median(tsteps.df$total.steps)
```

The mean of the total number of steps is
`r  round((mean(tsteps.df$total.steps)/mean(nsteps.per.day$total.steps)-1)*100,2)`
% greater after missing data imputation 
but the median has the same value. 


## Are there differences in activity patterns between weekdays and weekends?
